---
title: "API Example Catalog"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{API Example Catalog}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(plsSEMflow)
```

# Purpose

This vignette provides at least three short agronomic examples for the main public function families so that documentation is useful beyond the integrated tutorial.

# Learning sequence

1. State the agronomic question.
2. Identify the unit of observation and any grouping.
3. Define constructs before fitting.
4. Fit only an estimator that supports the requested model.
5. Diagnose measurement before structural interpretation.
6. Use user-controlled bootstrap uncertainty for inferential coefficients.
7. Assess prediction when prediction is claimed.
8. Export publication output without hiding the CI method or replication count.

# Agronomic example

```{r, eval=FALSE}
d <- pls_data("soil_crop")

m <- pls_model(
  pls_measurement(
    pls_reflective("SOIL", c("soil_om", "soil_cec", "soil_whc")),
    pls_reflective("NUTR", c("leaf_n", "leaf_p", "leaf_k")),
    pls_reflective("VIGOR", c("ndvi", "lai", "chlorophyll")),
    pls_reflective("YIELD", c("grain_yield", "biomass", "harvest_index"))
  ),
  pls_structural(
    pls_path("SOIL", "NUTR"),
    pls_path("NUTR", "VIGOR"),
    pls_path("VIGOR", "YIELD")
  )
)

fit <- pls_fit(m, d, engine = "native")
boot <- pls_bootstrap(fit, R = 5000, seed = 260819)
pls_table(fit, "paths", boot = boot)
pls_plot(fit, "paths", boot = boot)
```

# Interpretation discipline

Do not reduce the analysis to significance labels. Report the parameter estimate, uncertainty, construct meaning, unit and design context, and practical agronomic interpretation. When the analysis uses simulated teaching data, state that fact explicitly.

# What to verify before publication

- construct conceptualization;
- measurement quality;
- structural collinearity;
- bootstrap `R`, seed policy, CI method, and level;
- predictive assessment when relevant;
- invariance before group comparison;
- distinction between score-based sensitivity and dedicated latent estimators;
- reproducibility and package versions.

# Three additional examples

```{r, eval=FALSE}
# Example 1: bioinput composite
bio <- pls_data("bioinput_adoption")

# Example 2: ordinal farmer-adoption indicators
pa <- pls_data("precision_adoption")

# Example 3: clustered farms
ml <- pls_data("multilevel_farms")
```

# Connection to the integrated tutorial

See `21-foundations-to-advanced-tutorial.Rmd` for the full progression, publication examples, advanced backends, and reporting checklist.

# Module-specific scientific focus

**Teaching dataset:** `soil_crop`  
**Primary focus:** three examples per main API family.

The catalog complements roxygen help with agronomic examples at basic, intermediate and advanced levels.

The bundled dataset is frozen simulated teaching material. The workflow is realistic, but the resulting numerical values are not empirical evidence. In a real study, replace the teaching data while preserving the sequence of conceptualization, audit, estimation, uncertainty, diagnosis, interpretation, and reporting.


# Conceptual questions before running code

Before fitting any model in this module, answer the following questions in writing:

1. **What is the observational unit?** A plant, plot, farm, field, municipality, season, farmer, or repeated visit cannot be exchanged without changing the statistical interpretation.
2. **What exactly does each construct mean?** Give a scientific definition before looking at loadings or path coefficients.
3. **Why are the indicators attached to that construct?** State whether they are expected manifestations of a common factor or ingredients that jointly form a composite.
4. **Why is each arrow directional?** Use physiology, soil science, crop management, technology-adoption theory, chronology, or the experimental design. A directional arrow is not justified merely because a regression can be computed.
5. **Which claims are explanatory, predictive, or causal?** Keep these goals separate in the manuscript.
6. **Which uncertainty procedure will be reported?** For PLS-SEM coefficients, define bootstrap `R`, seed policy, confidence level, interval type, and resampling unit before producing the final table.

# Bootstrap layer

The examples in package help and evaluated vignettes can use a small replication count to keep checks fast. That value must not be copied mechanically into a final manuscript.

```{r, eval=FALSE}
# Fast code check only
boot_check <- pls_bootstrap(fit, R = 99, seed = 260819)

# Example of an analysis-scale setting selected by the analyst
boot_final <- pls_bootstrap(
  fit,
  R = 5000,
  seed = 260819,
  level = 0.95,
  interval = "percentile"
)
```

If greater Monte Carlo stability is required, increase `R`. The package does not hide the value or overwrite it with a fixed universal recommendation.

# Publication output

The same bootstrap object should be reused across tables and figures so that the reported uncertainty is internally consistent.

```{r, eval=FALSE}
path_tab <- pls_table(fit, "paths", boot = boot_final)
loading_tab <- pls_table(fit, "loadings", boot = boot_final)

p_path <- pls_plot(fit, "paths", boot = boot_final)
p_coef <- pls_plot(fit, "bootstrap", boot = boot_final)
```

A figure caption should identify the confidence level, bootstrap interval type, and number of resamples. A table footnote should provide the same information. Do not show confidence bars without stating how they were calculated.

# Interpretation template

A defensible paragraph normally contains four elements:

1. the direction and magnitude of the estimated relation;
2. the bootstrap confidence interval and its substantive width;
3. the amount of explained or predictive variation when relevant;
4. an agronomic interpretation constrained by the design.

Avoid converting every coefficient into a binary “significant/non-significant” conclusion. A narrow interval around a small effect and a wide interval around a large point estimate tell very different scientific stories.

# Sensitivity and escalation

Escalate the model only when a simpler representation leaves a specific problem unresolved. Examples include ordinal indicators, heterogeneous groups, nonlinear response, multilevel sampling, endogeneity, or prediction failure. Record the reason for escalation in the audit trail.

```{r, eval=FALSE}
pls_capabilities()
pls_external_capabilities()
```

An unavailable optional backend should produce an informative error. The core R workflow remains usable without Python, Julia, or commercial software.

# Common errors to avoid

- choosing a construct type from the observed correlation matrix alone;
- deleting an indicator automatically because a heuristic threshold is missed;
- using a universal sample-size rule without power or sensitivity reasoning;
- reporting paths without bootstrap uncertainty;
- interpreting R-squared as causal evidence;
- claiming prediction without held-out assessment;
- comparing groups without checking measurement comparability;
- using row-level resampling for a clustered inferential problem without justification;
- treating an exploratory score regression as the primary latent-variable estimator;
- changing engines until the preferred p-value appears;
- omitting software versions, seeds, bootstrap settings, and data-processing decisions.

# Minimum reporting block

For this module, report at least:

- dataset and unit of observation;
- construct definitions and measurement type;
- structural hypotheses;
- engine and estimator;
- missing-data handling;
- measurement diagnostics relevant to the construct type;
- structural estimates;
- bootstrap `R`, seed, CI level, and method;
- prediction or sensitivity diagnostics when relevant;
- limitations on causal or out-of-sample interpretation;
- package and R versions.

# Extended public-API examples

The examples below deliberately reuse the same agronomic domains at different levels. They are concise API references, while the integrated tutorial explains the scientific reasoning in detail.

## Assessment and extraction

```{r api-assessment, eval=FALSE}
# Example 1: complete soil-crop assessment
pls_assess(fit)
# Example 2: focus on the measurement results
pls_assess(fit)$measurement
# Example 3: focus on the structural results
pls_assess(fit)$structural

# Example 1: inspect the native result list
pls_backend_object(fit)
# Example 2: inspect an optional cSEM object
pls_backend_object(fit_csem)
# Example 3: preserve the raw object for an audit bundle
saveRDS(pls_backend_object(fit), "native-backend-object.rds")

# Example 1: inspect scores for agronomic constructs
pls_scores(fit)
# Example 2: append scores for exploratory field diagnostics
cbind(d, pls_scores(fit))
# Example 3: use scores only in an explicitly labeled sensitivity analysis
summary(pls_scores(fit)$VIGOR)
```

## Construct and relation specification

```{r api-specification-extra, eval=FALSE}
# Example 1: soil composite
pls_composite("SOIL_CAP", c("soil_om", "soil_cec", "soil_whc"))
# Example 2: management composite with Mode A as a sensitivity specification
pls_composite("MGT", c("management_planning", "input_timing", "records_quality"), mode="A")
# Example 3: bioinput-management composite
pls_composite("CAPACITY", c("training_hours", "technical_assistance_visits", "credit_access"))

# Example 1: soil -> nutrition -> yield mediation
pls_mediation("SOIL", "NUTR", "YIELD")
# Example 2: usefulness -> intention -> use
pls_mediation("PU", "INTENT", "USE")
# Example 3: management -> vigor -> production
pls_mediation("MGT", "VIGOR", "PROD", direct=TRUE)

# Example 1: drought moderation of management effect
pls_moderation("MGT", "DROUGHT", "STABILITY", method="two_stage")
# Example 2: soil condition as moderator
pls_moderation("NUTR", "SOIL", "VIGOR", method="orthogonal")
# Example 3: technology readiness as moderator of usefulness
pls_moderation("PU", "READINESS", "INTENT", method="product_indicator")

# Example 1: quadratic irrigation response
pls_quadratic("IRRIGATION", "EFFICIENCY")
# Example 2: quadratic stress response
pls_quadratic("STRESS", "STABILITY")
# Example 3: quadratic input-intensity response
pls_quadratic("INPUT", "YIELD")

# Example 1: degree-2 sensitivity
pls_nonlinear("STRESS", "STABILITY", degree=2)
# Example 2: cubic response exploration
pls_nonlinear("MGT", "EFFICIENCY", degree=3)
# Example 3: nonlinear adoption response
pls_nonlinear("READINESS", "USE", degree=2)

# Example 1: farms nested in municipality
pls_cluster("municipality")
# Example 2: repeated farms within region when region is the inferential cluster
pls_cluster("region")
# Example 3: fields nested in producer
pls_cluster("producer_id")
```

## Effects, moderation, nonlinear response, and groups

```{r api-effects-extra, eval=FALSE}
# Example 1: direct and mediated effects
pls_effects(fit)
# Example 2: retain decomposition for a manuscript table
effects_soil <- pls_effects(fit)
# Example 3: inspect only indirect effects
subset(pls_effects(fit), grepl("indirect", effect))

# Example 1: drought moderation
pls_moderation_test(fit_irrig, "MGT", "STRESS", "STABILITY", R=1999, seed=260819)
# Example 2: more bootstrap resamples for final reporting
pls_moderation_test(fit_irrig, "MGT", "STRESS", "STABILITY", R=10000, seed=260819)
# Example 3: point estimates only for a rapid diagnostic
pls_moderation_test(fit_irrig, "MGT", "STRESS", "STABILITY", boot=FALSE)

# Example 1: quadratic score sensitivity
pls_nonlinear_test(fit_irrig, "STRESS", "STABILITY", degree=2, R=1999, seed=260819)
# Example 2: cubic sensitivity
pls_nonlinear_test(fit_irrig, "STRESS", "STABILITY", degree=3, R=1999, seed=260819)
# Example 3: fast point-estimate screen
pls_nonlinear_test(fit_irrig, "STRESS", "STABILITY", degree=2, boot=FALSE)

# Example 1: irrigated versus rainfed
pls_mga(m, d, group="system", R=1999, seed=260819)
# Example 2: increase R for final inference
pls_mga(m, d, group="system", R=10000, seed=260819)
# Example 3: save comparison for reporting
mga_soil <- pls_mga(m, d, group="system", R=4999, seed=260819)
```

## cSEM specialist post-estimation

```{r api-csem-extra, eval=FALSE}
fit_csem <- pls_fit(m, d, engine="cSEM", estimator="PLSc")

# Three assessment uses
pls_csem_assess(fit_csem)
pls_csem_assess(fit_csem, .quality_criterion="all")
assessment_csem <- pls_csem_assess(fit_csem)

# MICOM with progressively stronger resampling
pls_micom(fit_group, R=999, seed=260819)
pls_micom(fit_group, R=4999, seed=260819)
pls_micom(fit_group, R=10000, seed=260819)

# Multigroup-difference testing
pls_mgd(fit_group)
mgd_result <- pls_mgd(fit_group)
print(pls_mgd(fit_group))

# Endogeneity sensitivity
pls_endogeneity(fit_csem)
endog <- pls_endogeneity(fit_csem)
print(pls_endogeneity(fit_csem))

# IPMA
pls_ipma(fit_csem)
ipma <- pls_ipma(fit_csem)
print(pls_ipma(fit_csem))

# Dedicated nonlinear-effects backend
pls_nonlinear_effects(fit_nl_csem)
nl_effects <- pls_nonlinear_effects(fit_nl_csem)
print(pls_nonlinear_effects(fit_nl_csem))
```

## Engine comparison and external validation

```{r api-integration-extra, eval=FALSE}
# Compare available R engines
pls_compare_engines(m, d, engines=c("native","cSEM"))
pls_compare_engines(m, d, engines=c("native","seminr"))
engine_check <- pls_compare_engines(m, d, engines=c("native","cSEM","plspm"))

# Export three agronomic cases for external validation
pls_export_external(m, d, dir="external-soil-crop")
pls_export_external(m_irrig, irrig, dir="external-irrigation")
pls_export_external(m_adopt, pa, dir="external-precision-adoption")

# Import result tables from specialist software
pls_import_external("external-soil-crop/results.csv")
pls_import_external("external-irrigation/results.csv")
external_paths <- pls_import_external("external-soil-crop/results.csv")

# Compare independent external path estimates
pls_compare_external(fit, external_paths)
external_comparison <- pls_compare_external(fit, external_paths)
print(pls_compare_external(fit, external_paths))
```

## Python and Julia bridges remain optional

```{r api-bridges-extra, eval=FALSE}
# Python SEMinR port: soil-crop model
pls_python_seminr(m, d)
# Python bridge on an adoption model
pls_python_seminr(m_adopt, pa)
# Explicit Python environment selection
pls_python_seminr(m, d, envname="plssem-python")

# Julia companion bundle for SEM benchmarking
pls_julia_bundle(m, d, dir="julia-soil-crop")
pls_julia_bundle(m_adopt, pa, dir="julia-adoption")
pls_julia_bundle(m, d, dir="julia-reproducibility-check")
```

## Heterogeneity

```{r api-heterogeneity-extra, eval=FALSE}
fit_plspm <- pls_fit(m, d, engine="plspm")

# REBUS with default convergence criterion
pls_rebus(fit_plspm)
# REBUS with stricter convergence
pls_rebus(fit_plspm, stop_crit=.001, iter_max=200)
# Preserve the heterogeneity object
rebus_soil <- pls_rebus(fit_plspm, stop_crit=.005)

# External specialist workflows
pls_heterogeneity(fit, "FIMIX")
pls_heterogeneity(fit, "PLS-POS")
pls_heterogeneity(fit_plspm, "REBUS")
```

## Publication output

```{r api-publication-extra, eval=FALSE}
# Coefficient forest from moderation
mod <- pls_moderation_test(fit_irrig, "MGT", "STRESS", "STABILITY", R=4999, seed=260819)
pls_plot_coefficients(mod, title="Drought-stress moderation")
pls_plot_coefficients(mod, file="moderation.pdf")
pls_plot_coefficients(mod, file="moderation.tiff", dpi=600)

# MGA plot
mga <- pls_mga(m, d, "system", R=4999, seed=260819)
pls_plot_mga(mga)
pls_plot_mga(mga, file="irrigation-system-mga.pdf")
pls_plot_mga(mga, file="irrigation-system-mga.tiff", dpi=600)

# Advanced tables
pls_table_advanced(mod)
pls_table_advanced(mod, format="markdown")
pls_table_advanced(mga, format="flextable")

# Excel export for three result classes
pls_export_xlsx(pls_table(fit,"paths",boot=boot_final), "paths.xlsx")
pls_export_xlsx(pls_table(fit,"loadings",boot=boot_final), "loadings.xlsx")
pls_export_xlsx(pls_table_advanced(mod), "moderation.xlsx")

# Reproducible reports
pls_report(fit, "soil-crop-report.md", boot=boot_final)
pls_report(fit_irrig, "irrigation-report.md")
pls_report(fit, "soil-crop-reviewer-audit.md", title="PLS-SEM reviewer audit", boot=boot_final)
```

## Simulation and validation

```{r api-validation-extra, eval=FALSE}
truth <- c("SOIL -> NUTR"=.50, "NUTR -> VIGOR"=.40, "VIGOR -> YIELD"=.45)

# Three sample-size scenarios
pls_simulate(m_simple, n=200, path_values=truth, seed=260819)
pls_simulate(m_simple, n=500, path_values=truth, seed=260820)
pls_simulate(m_simple, n=1000, path_values=truth, seed=260821)

# Validation cases with user-selected bootstrap R
pls_validation_case(m_simple, n=300, path_values=truth, R=499, seed=260819)
pls_validation_case(m_simple, n=600, path_values=truth, R=1999, seed=260819)
pls_validation_case(m_simple, n=1200, path_values=truth, R=4999, seed=260819)

# Cross-engine numerical agreement under three tolerances/scenarios
pls_validate_cross_engine(m, d, engines=c("native","cSEM"), tolerance=.10)
pls_validate_cross_engine(m, d, engines=c("native","cSEM"), tolerance=.05)
pls_validate_cross_engine(m_simple, sim, engines=c("native","cSEM"), tolerance=.03)
```

## Dataset registry and capability inspection

```{r api-registry-extra, eval=FALSE}
# Dataset registry in three teaching contexts
pls_datasets()
subset(pls_datasets(), grepl("ordinal", primary_use))
subset(pls_datasets(), grepl("multilevel", primary_use))

# Backends and capabilities
pls_capabilities()
subset(pls_capabilities(), installed)
subset(pls_capabilities(), tier <= 2)
```
